---
title: "myscript"
author: "Taylor PD"
date: "2025-01-07"
output:
  html_document: default
  pdf_document: default
---
[GitHub Repository](https://github.com/Taylorpd4/BIOL432_Assignment1TP)

## Introduction
This report shows the introductory process including the necessary skills that are applyed when analyzing biological measurement data. It executes two R scripts, `dataGenerato.R` and `volumeEstimato.R`, to create and enhance the dataset. This report also includes data sorting, summary statistics, and visualizations.
```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Step 1: Execute R Scripts
### 1.1 Generate Data
The script `dataGenerato.R` creates a dataset with 100 observations, including species names, limb measurements, and observer names. the dataset is saved as `measurements.csv`.

```{r}
source("dataGenerato.R")
```

### 1.2 Estimate Volume
The script `volumeEstimato.R` calculates the limb volume for each observation. It adds this data as a new column to `measurements.csv`.

```{r}
source("volumeEstimato.R")
```

## Step 2: Load and Prepare Data

```{r}
data <- read.csv("measurements.csv")
```

### 2.1 Sort Data by Species, Observer, and Limb Volume

```{r}
sorted_data <- data %>% arrange(Species, Observer, Volume)
head(sorted_data)
```

### 2.2 Create Table of Average Volume by Species

```{r}
average_volume <- data %>% 
  group_by(Species) %>% 
  summarise(Average_Volume = mean(Volume, na.rm = TRUE))

knitr::kable(average_volume, caption = "Average Limb Volume by Species")
```

### 2.3 Create Table of Observation Counts by Species and Observer

```{r}
observation_counts <- data %>% 
  group_by(Species, Observer) %>% 
  summarise(Count = n(), .groups = "drop")

knitr::kable(observation_counts, caption = "Observation Counts by Species and Observer")
```

## Step 3: Visualizations

### 3.1 Box Plot of Volume by Species

```{r}
library(ggplot2)
ggplot(data, aes(x = Species, y = Volume, fill = Species)) + 
  geom_boxplot() + 
  labs(title = "Distribution of Limb Volumes by Species", x = "Species", y = "Limb Volume") + 
  theme_minimal()
```

### 3.2 Multi-Panel Histograms of Volume by Species

```{r}
ggplot(data, aes(x = Volume)) + 
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) + 
  facet_wrap(~ Species, scales = "free") + 
  labs(title = "Frequency Distribution of Limb Volume by Species", x = "Limb Volume", y = "Frequency") + 
  theme_minimal()
```

## Conclusion
This report provides a complete workflow for analyzing biological measurement data.


```



